### 2017-10-22 paper01_peloquin_frank2016/model_walkthru/rsa_sandbox.rmd
fix a few things start playing around

---
title: "rrrsa imp"
output: html_document
---

```{r setup}
lefftpack::lazy_setup()
```

```{r normalize}
normalize <- function(vec){
  # all elements of `vec` must be positive or zero 
  stopifnot(min(vec, na.rm=TRUE) >= 0)
  # if `vec` is all zeros, just return it right back 
  if (length(unique(vec))==1){
    if (unique(vec)==0){
      return(vec)
    }
  }
  # normalize by dividing each element by the sum 
  return(vec / sum(vec, na.rm=TRUE))
}
```

```{r normalize_matrix}
normalize_matrix <- function(mat, rows_or_cols){
  if (!rows_or_cols %in% c("rows", "cols")){
    message("gotta normalize over rows or cols mayne!")
    return(NULL)
  }
  if (rows_or_cols=="rows"){
    return(t(apply(mat, MARGIN=1, normalize)))
  } 
  if (rows_or_cols=="cols"){
    return(apply(mat, MARGIN=2, normalize))
  }
}
```

```{r utility}
utility <- function(items, costs, alpha){
  # `items` is a vector of nums, each representing a word/item 
  # `costs` must have same length as `items`, `alpha` a positive scalar

  # apply `informativity()` to each item with specified costs and alpha. 
  # then collect them all and return with the same shape/order as `items`
  inform_unnorm <- mapply(informativity, prob=items, cost=costs, alpha=alpha)

  # normalization here is just dividing each number by the sum of them all 
  inform_norm <- normalize(inform_unnorm)

  return(inform_norm)
}

```

```{r informativity}
informativity <- function(prob, cost, alpha){
  # `prob` must be a probability, `alpha` a positive scalar 
  inform <- ifelse(prob == 0, 0, exp(-alpha * (-log(prob) - cost)))
  return(inform)
}

```

```{r full_recursion}
full_recursion <- function(mat, costs, priors, alpha){

  cost_mat <- matrix(
    nrow=nrow(mat), ncol=ncol(mat), 
    dimnames=dimnames(mat), 
    data=rep(as.numeric(costs), times=1, each=nrow(mat))
  )

  # the likelihood function is `utility()` 
  likelihood <- t(sapply(1:nrow(mat), function(row){
    utility(items=mat[row, ], costs=cost_mat[row, ], alpha=alpha)
  }))

  unnorm_posterior <- apply(likelihood, MARGIN=2, function(col) priors * col)

  posterior <- normalize_matrix(unnorm_posterior, rows_or_cols="cols")

  dimnames(posterior) <- list(rownames(mat), colnames(mat))

  return(posterior)
}

```

```{r rsa_convenience}
rsa_matrix_to_df <- function(rsa_matrix, response_colname){
  out <- as.data.frame(rsa_matrix) %>% 
    mutate(quantity = as.numeric(gsub("row", "", rownames(rsa_matrix)))) %>% 
    melt(id.vars="quantity", variable.name="word") %>% 
    mutate(word = as.character(word)) %>% 
    arrange(quantity, word) 
    # rename(semantics = "value")

  names(out)[names(out)=="value"] <- response_colname

  return(out)
}

make_rsa_plot <- function(input_matrix, output_matrix, facet_words=TRUE){

  input_df <- rsa_matrix_to_df(input_matrix, response_colname="semantics")
  output_df <- rsa_matrix_to_df(output_matrix, response_colname="pragmatics")

  input_output <- 
    full_join(input_df, output_df, by=c("quantity", "word")) %>% 
    melt(id.vars=c("quantity", "word")) %>% 
    rename(type = variable) %>% 
    mutate(type = as.character(type))

  input_output$type <- factor(input_output$type, 
                              levels=c("semantics","pragmatics"))

  if (facet_words){
    the_plot <- 
      ggplot(input_output, aes(x=quantity, y=value, color=type)) + 
      geom_point(size=rel(.5)) + 
      geom_line() + 
      scale_color_manual(values=c("blue", "orange")) + 
      facet_wrap(~word)

  } else {
    the_plot <- 
      ggplot(input_output, aes(x=quantity, y=value, color=word)) + 
      geom_line() + 
      facet_wrap(~type)
  }

  return(the_plot)
}

rsaplot <- function(smat, costs, priors, alpha, facet_words=TRUE){
  if (!all(colSums(smat)==1)){
    smat <- normalize_matrix(smat, rows_or_cols="cols")
  }
  if (sum(priors)!=1){
    priors <- normalize(priors)
  }
  pmat <- full_recursion(smat, costs, priors, alpha)
  return(make_rsa_plot(smat, pmat, facet_words=facet_words))
}

```

```{r semantics_matrix}
words <- c("none","none_all","some","some_notall","most","all","not_all")
costs <- setNames(object=c(1, 1, 1, 1, 1, 1, 1), nm=words)

points <- 1:5
priors <- normalize(rep(1, length(points)))

alpha <- 1

semantics_unnorm <- matrix(
  data=rep(NA, length(points)*length(words)), 
  nrow=length(points), ncol=length(words), 
  dimnames=list(paste0("row", points), words)
)
semantics_unnorm[, "none"]        <- c(1, 0, 0, 0, 0)
semantics_unnorm[, "none_all"]    <- c(1, 0, 0, 0, 1)
semantics_unnorm[, "some"]        <- c(0, 1, 1, 1, 1)
semantics_unnorm[, "some_notall"] <- c(0, 1, 1, 1, 0)
semantics_unnorm[, "most"]        <- c(0, 0, 1, 1, 1)
semantics_unnorm[, "all"]         <- c(0, 0, 0, 0, 1)
semantics_unnorm[, "not_all"]     <- c(1, 1, 1, 1, 0)
semantics_unnorm

rsaplot(semantics_unnorm, costs, priors, alpha)

s_a_na <- matrix(
  data=c(0,1,1,1,1,   0,0,0,0,1,    1,1,1,1,0), 
  nrow=5, ncol=3, dimnames=list(paste0("row", 1:5), 
                                c("some","all","not_all"))
)
rsaplot(s_a_na, costs=c(1,1,1), priors=rep(1, 5), alpha=1)

s_a <- matrix(
  data=c(0,1,1,1,1,   0,0,0,0,1), 
  nrow=5, ncol=2, dimnames=list(paste0("row", 1:5), 
                                c("some","all"))
)
rsaplot(s_a, costs=c(1,1), priors=rep(1, 5), alpha=1)

s_na <- matrix(
  data=c(0,1,1,1,1,   1,1,1,1,0), 
  nrow=5, ncol=2, dimnames=list(paste0("row", 1:5), 
                                c("some","not_all"))
)
rsaplot(s_na, costs=c(1,1), priors=rep(1, 5), alpha=1)

```