### 2017-11-03 rsa-math.rmd
put up materials for LG15 and FG12 nov03

---
title: ""
output: pdf_document
font_size: 10pt
geometry: margin=.25in
---

## RSA math cheatsheet (material from Frank and Goodman 2012 -- see paper for context)

#### Equation 1 (FG2012) -- posterior probability

$$P(r_s|w, C) = \frac{P(w|r_s, C)\times P(r_s)}{\sum\limits_{r'\in C} P(w|r', C) \times P(r')}$$

*(i)* $r_s$ is an intended referent (element of $C$); *(ii)* $C$ is a context (set of objects $r, r', ...\in C$); *(iii)* $w$ is the word uttered by the speaker; *(iv)* the **prior** is $P(r_s)$; *(v)* the **likelihood** is $P(w|r_s, C)$; and *(vi)* the **normalizing constant** is $\sum\limits_{r'\in C} P(w|r', C) \times P(r')$

#### Equation 2 (FG2012) -- likelihood for FG12 (derived from formulas below)

$$P(w|r_s, C) = \frac{|w|^{-1}}{\sum\limits_{w'\in W} |w'|^{-1}}$$

- $|w|$ is the number of objects in $C$ to which word $w$ could(?) apply, 
- and $W$ is the set of words that apply to the intended referent $r_s$.

#### Equation S1 (FG12 supplement)

$$P(w|r_s, C) \propto e^{\alpha\times U(w; r_s, C)}$$

*(i)* the **context** $C = \{o_1, ..., o_n\}$ is a set of objects; *(ii)* the **vocabulary** is a set of words $V = \{w_1, ..., w_m\}$; *(iii)* each word $w\in V$ has a meaning $[\![ w ]\!] : C \rightarrow \{0, 1\}$; and *(iv)* words are chosen in proportion to their expected utility

#### Equation S2 -- utility (FG12 supplement)

$$U(w; r_s, C) = I(w; r_s, C) - D(w)$$

where $I(w; r_s, C)$ is the informativeness of utterance $w$ w.r.t. intended referent $r_s \in C$; and $D(w)$ is $w$'s cost (in $C$?).

<br>

#### Surprisal (FG12 supplement), for a sample $x$ from a known distribution $p(x)$ 
$$I_p(x) = -log(p(x))$$

<br>

#### Utility and surprisal (FG12 supplement) -- utility decreases with surprisal

$$I(w; r_s, C) = -I_{\widetilde{w}_C}(r_s)$$

where $\widetilde{w}_C$ is the distribution over objects that would come froma  aliteral interpretation of $w$ in context $C$.

<br>

#### Equation S3 -- literal listener model derivation (FG12, supplementary materials)

"if listeners interpret the utterance $w$ literally, assigning zero probability to objects for which the word is false, they assign equal probability to each object conssitent with $w$. This distribution over objects can be written:"

$$\widetilde{w}_C(o) = \frac{1}{|w|} \textrm{ if } w(o) = true \textrm{, and otherwise } 0$$

<!-- $$\widetilde{w}_C(o) =  -->
<!--     \begin{cases} -->
<!--     \frac{1}{|w|} & \text{if } w(o) = true\\\\ -->
<!--     0             & \text{otherwise } -->
<!--     \end{cases}$$ -->

<br>

#### Equation S4 -- listener likelihood (FG12, supplement)

equation S4 follows from S1-S3, which is equivalent to equation 2 (aka "the size principle")

$$P(w|r_s, C) = \frac
                    {e^{-(-\log(|w|^{-1}))}}
                    {\sum\limits_{w'\in V \text{s.t. } w'(r_s) = true} e^{-(-\log(|w'|^{-1}))}}$$

<style>
<!-- /* use this to get font import code: http://fonts.googleapis.com/css?family=FONT+NAME*/ -->
<!-- /* use **latin**, not latin-extended */ -->
@font-face {
  font-family: 'Source Sans Pro';
  font-style: normal;
  font-weight: 400;
  src: local('Source Sans Pro Regular'), local('SourceSansPro-Regular'), url(https://fonts.gstatic.com/s/sourcesanspro/v11/ODelI1aHBYDBqgeIAH2zlJbPFduIYtoLzwST68uhz_Y.woff2) format('woff2');
  unicode-range: U+0000-00FF, U+0131, U+0152-0153, U+02C6, U+02DA, U+02DC, U+2000-206F, U+2074, U+20AC, U+2212, U+2215;
}

@font-face {
  font-family: 'Roboto Mono';
  font-style: normal;
  font-weight: 400;
  src: local('Roboto Mono'), local('RobotoMono-Regular'), url(https://fonts.gstatic.com/s/robotomono/v5/hMqPNLsu_dywMa4C_DEpY4gp9Q8gbYrhqGlRav_IXfk.woff2) format('woff2');
  unicode-range: U+0000-00FF, U+0131, U+0152-0153, U+02C6, U+02DA, U+02DC, U+2000-206F, U+2074, U+20AC, U+2212, U+2215;
}

body {
  padding: 10px;
  font-size: 12pt;
  font-family: 'Source Sans Pro', sans-serif;
}

h4 {
  margin-top: 32px;
  font-weight: bold;
}

blockquote {
  font-size: 12pt;
  font-family: 'Source Sans Pro', sans-serif;
}

code {
  font-family: 'Roboto Mono', monospace;
  font-size: 13px;
}

pre {
  font-family: 'Roboto Mono', monospace;
  font-size: 13px;
}

.table {
  width: 75%;
  align: center;
}

h3.subtitle {
  font-size: 26px;
  /*color: #aca39a;*/
  font-style: normal !important;
  font-weight: lighter !important;
  border-bottom: none;
  padding-top: 4px !important;
}

h4.author { 
  font-size: 22px;
  color: #aca39a;
  font-weight: lighter !important;
  padding-top: 0px !important;
  margin-top: 2px !important;
  margin-bottom: 12px !important;
  padding-bottom: 12px !important;
}

h3 {
  margin-top: 42px;
}

h2 {
  margin-top: 48px;
  border-bottom: 2px solid; 
}

</style>